Direct observation of \swave scattering of fermions crossing a Feshbach resonance

### 2015-03-25 draft1.tex
2nd draft given to Ian.

\begin{document}
\newcommand{\swave}[0]{$\it{s}$-wave }
\newcommand{\pwave}[0]{$\it{p}$-wave }
\newcommand{\K}{ $^{40}\rm{K}$}
\newcommand{\us}{$\mu \rm{s}$}

@@

\section{S-wave scattering experiment}
In this section we describe our experiment with scattering two counter-propagating clouds of $^{40}K$ atoms and observe the resulting s-wave halo of scattered atoms. We vary the magnetic field around the Feshbach resonance during this experiment, and from the change in scattering deduce a location of the magnetic fields resonance of 202.47 Gauss and a width of 8.98 Gauss.
\subsection{Experimental procedure}
We load our cold Fermi gas of $^{40}K$ atoms into a crossed dipole trap. We then use adiabatic rapid passage to bring the gas into an equal superposition of $\ket{9/2,-9/2}$ and $\ket{9/2,-7/2}$, the states whose Feshbach resonance we are trying to study. 
\par We then use a large pair of  coils in Helmholtz configuration to bring the magnetic field close to the Feshbach resonance, but not close enough for significant collisions to occur. We hold the atoms there to allow the eddy currents in the large coils to settle, and then use a smaller set of Helmholtz coils to hop the field to the desired value. We take two sets of data: one coming from below the resonance, where we hop the large coils to a lower field value than the resonance and use the small coils to hop up to the desired field, and one coming from below the resonance, where we hop the large coils to a field value above the resonance and then use the small coils to hop the field down to the desired value. This allows us to correct for losses due to molecule formation and three-body recombination \cite{Chin10}.
\par We then use a double pulse sequence \cite{Wu05} of a near resonant 1-d retro-reflected optical lattice to transfer most of the atoms into the first two excited states of the lattice, giving them $\pm 2k_r$ of momentum, where $k_r$ is the recoil momentum of the lattice. Since the initial Fermi gas has a wide momentum spread (in contrast to a BEC, which has a very narrow momentum spread), and the lattice pulsing is a momentum dependent process, not all the atoms are successfully transferred into the 1st excited band of the lattice. We optimized our pulse times to minimize the number of atoms that remain in the lowest, zero momentum, band of the lattice, but were not able to eliminate them completely. The optimized pulse times are 23\us{} for the first square pulse, 13\us{} off interval, and 12\us{} for the second square pulse. 
\par We release the atoms from the trap and allow 1ms for the two opposite momentum states within the cloud to pass through each other, scattering on the way. For the data taken coming from below the Feshbach resonance, we then simply ramp down the field and image the atoms. For the data taken coming from above the Feshbach resonance, we then ramp the field back up through the resonance to recover any molecules that were created when the field was ramped from the attractive to the repulsive side of the resonance, and then quickly ramp the field back down and image the atoms. We use an imaging pulse of 40\us{}.

@@

We first correct each image for recoil induced detuning using the simulation described above. The simulation gives us a look-up table of observed optical density as a function of atom column density and probe intensity. From our two absorption images we observe the probe intensity and optical density for each pixel. We then use the simulation look-up table in an inverted faction to deduce the column density or, equivalently, the optical depth. An example of the effect of this correction procedure in shown in figure \ref{fig:SampleCorrection}

@@

\caption{An example of our absorption image after 6.784ms time of flight. The 1-D lattice is along the horizontal direction at the center of our images. The two large clouds on the left and right are the atoms in the $\pm 2k_r$ momentum orders that passed through each other unscattered. The smaller cloud in the center is the atoms that remained in the lowest band of the lattice after pulsing, and thus obtained no momentum. The thin spread of atoms around these clouds is the atoms that underwent scattering.  a. Raw optical density b. Optical density of the same image after correcting for recoil induced detuning by comparing to simulation}  

@@

\par To improve the signal and compensate for our shot to shot number fluctuations, we take 15 equivalent images for each data point and, after correcting, add them together. We then perform a standard inverse Abel transform to extract the radial atom number dependence from the two dimensional projection we see in our images. The inverse Abel transform assumes a cylindrical symmetry, which is present in our case, with the axis of symmetry defined by the 1-D lattice. 
\par We then extract the number of atoms that went through a single scattering event $N_{scat}$, as a fraction of the total atom number $N_{tot}$, for each bias magnetic field value. We unscattered number by counting the number of atoms in the two unscattered clouds. We obtain the number of atoms that underwent single scattering by counting the number of atoms outside the Fermi radius of the unscattered clouds, but inside the arc created by rotating the Fermi radius around the original center of the cloud. This method excludes atoms that underwent multiple scatterings, as those atoms are not bound by momentum conservation to have traveled the same distance from the center as the unscattered clouds in time of flight. The total atom number is the sum of those two.
\par We then use our data to deduce the resonant field value, the width of the resonance, etc. To do this, we must fit our data to a model. From \cite{Chin10}, we have the expression:
a(B)=a_{bg}\left(1-\frac{\Delta}{B-B_0}\right),
where $a$ is the scattering length, $a_{bg}$ is the scattering lenth away from the resonance, $\Delta$ is the width of the resonance, and $B_0$ is the field value at which the resonance occurs. Since we are in the low energy regime, the scattering cross-section is given by $\sigma=4\pi a^2$. 

@@

\frac{N_{scat}}{N_{tot}^2}=\tilde{a}_{bg}^2\left(1-\frac{\Delta}{B-B_0}\right)^2.

@@

\par Unfortunately, we cannot fit to this equation directly - camera readout noise is skewed towards the positive, creating an asymmetry in detected shot noise for very low and zero atom number pixels. Thus, there is a background number of what looks like scattered atoms in our treatment, giving rise to a background offset $o$ in our data. Thus, the equation we actually fit to is
\frac{N_{scat}}{N_{tot}^2}=\tilde{a}_{bg}^2\left(1-\frac{\Delta}{B-B_0}\right)^2 + o,
\label{eq13}
where our fit parameters are $\tilde{a}_{bg}$, $\Delta$, $B_0$, and $o$. 

@@

Our final data is presented in figure \ref{fig:fittedFractions}. The red line is a best fit of the model given in eq.\ref{eq13}. The fit parameters we extract are $\tilde{a}_{bg}=1.48e-3$, $\Delta = 8.98$ Gauss, $B_0 = 202.47$ Gauss, and $o=9.00e-5$. The error bars on the fitted data are obtained solely from photon shot noise of both absorption images propagated through our analysis.
The accepted values for the $^{40}K$ s-wave Feshbach resonance for the  $\ket{9/2,-9/2}$ and $\ket{9/2,-7/2}$ states are $B_0=202.1\pm0.07$ and $\Delta=7.8\pm0.6$. 

@@

\caption{Normalized scattered population plotted versus bias field B. Green dots represent data taken coming from below the resonance, and blue dots represent the data taken coming from above the resonance. Red is the line of best fit, where data coming from above the resonance was used above the resonance and data coming from below the resonance was used below the resonance to create the fit. The regime where the scattering length is large enough for the atoms to behave hydrodynamically is shaded in gray. Data points in that region were not used in the fit, as there the assumption $\sigma\rho\ll1$, where $\rho$ is the atom number per unit area, is no longer valid. Values for the resonant field $B_0$ from literature and as found in this work are indicated.    }  

@@

We studied the effects of recoil-induced detuning effects on absorption images and found an optimal imaging time of ~40\us{} for $^{40}K$ atoms for noise minimization after corrections. We use these results to observe s-wave scattering halos of the Fermi gas around the Feshbach resonance and directly verify the resonance location and width. Our analysis can be used in any absorption imaging applications where signal to noise minimization is critical. Experimental techniques developed for our s-wave scattering experiment can be readily extended to a Raman-coupled single component gas version, in which we hope to detect synthetic p-wave interactions. 

>>>>

\begin{document}
\newcommand{\swave}[0]{$\it{s}$-wave }
\newcommand{\pwave}[0]{$\it{p}$-wave }
\newcommand{\K}{ $^{40}\rm{K}$}
\newcommand{\Rb}{ $^{87}\rm{Rb}$}
\newcommand{\us}{$\mu \rm{s}$}

@@

\section{S-wave scattering experiment}
In this section we describe our Fermi scattering experiment. We scattered two counter-propagating clouds of $^{40}K$ atoms and observed the resulting \swave halo of scattered atoms.  We measured the dependence of the scattered atomic fraction on the bias magnetic field in the vicinity of the Feshbach resonance. We used this data to extract the location of the magnetic fields resonance of 20.247(2) mT and a width of 1.0(1) mT.
\subsection{Experimental procedure}
Our cold Fermi cloud preparation was as follows. We prepared clouds of cold \K{} atoms in a hybrid \K{} and \Rb{} apparatus. We used a Zeeman slower to slow both species before being captured in a Magneto-Optical Trap (MOT). We allowed both species to cool in optical molasses for 2 miliseconds. Then, the \K{} atoms were optically pumped into the magnetically trappable $\ket{F=9/2, m_F=9/2}$ state. Both species were then loaded into a quadrupole magnetic trap, and cooled evaporatively. Since the \K{} atoms only interact very weakly with each other, they only re-thermalize due to interaction with \Rb{} atoms, and therefore the \Rb{} atoms are necessary to evaporatively cool \K{}. We then load the atoms into a crossed optical dipole trap, with trap frequencies of 39, 42, 124 Hz in the three spatial directions [I GOT THOSE FROM ROSS'S PAPER COULD BE WRONG FOR US]. We continued evaporative cooling by slowly ramping down the dipole trap. We then used adiabatic rapid passage (ARP) to transfer the \Rb{} atoms from the $\ket{F=2, m_F=2}$ state to the  $\ket{F=1, m_F=-1}$ absolute ground state. This state was chosen to minimize spin changing collisions with \K{} atoms during any further evaporation.  We then briefly shined an on-resonant probe light to force any remaining $F=2$ atoms out of the trap. We again used ARP to transfer the \K{} atoms into an equal superposition of $\ket{F=9/2,m_F=-9/2}$ and $\ket{F=9/2,m_F=-7/2}$, and further evaporated in the dipole trap. The $\ket{9/2,-9/2}$ and $\ket{9/2,-7/2}$ hyperfine states of \K{} were then used to study their Feshbach resonance.

\par We then ramped the bias field in a two-step fashion to the desired value $B$ near the Feshbach resonance. To approach the field, we used a large pair of  coils in Helmholtz configuration to bring the magnetic field 0.59 mT of $B$. We held the atoms there to allow the eddy currents in the large coils to settle, and then used a smaller set of Helmholtz coils to hop the field the remaining 0.59 mT. We took two sets of data: one coming from below the resonance, where we hopped the large coils to $B-0.59 mT$ and used the small coils to hop up to the desired field, and one coming from above the resonance, where we hopped the large coils to  $B+0.59 mT$ and then used the small coils to hop the field down to the desired value. This allowed us to correct for losses due to molecule formation and three-body recombination \cite{Chin10}.
\par Once we had the Fermi cloud at the intended bias field, we imparted momentum in the $x$ direction to half of the cloud and in the $-x$ direction to the other half and observed them scatter as they moved through each other and separated. To impart the momentum, we used a double pulse sequence \cite{Wu05} of a near resonant 1-d retro-reflected optical lattice. The pulse sequence was optimized to transfer most of the atoms into the first two excited states of the lattice, giving them $\pm 2k_r$ of momentum, where $k_r$ is the recoil momentum of the lattice. Since the initial Fermi gas had a wide momentum spread (in contrast to a BEC, which has a very narrow momentum spread), and the lattice pulsing is a momentum dependent process, not all the atoms were successfully transferred into the 1st excited band of the lattice. We optimized our pulse times to minimize the number of atoms that remain in the zero momentum band of the lattice, but were not able to eliminate them completely. The optimized pulse times are 23\us{} for the first square pulse, 13\us{} off interval, and 12\us{} for the second square pulse. 
\par We then released the atoms from the trap and allowed 1ms for the two opposite momentum states within the cloud to pass through each other, scattering on the way. For the data taken coming from below the Feshbach resonance, we then simply ramped down the field and imaged the atoms. For the data taken coming from above the Feshbach resonance, we ramped the field back up through the resonance to recover any molecules that were created when the field was ramped from the attractive to the repulsive side of the resonance, and then quickly ramped the field back down and imaged the atoms. We use an imaging pulse of 40\us{}. 
\par The total time from the moment the atoms were released from the trap to when they were imaged, known as the time of flight, was 6.784 ms. In such an image, the location of each atoms is determined by how far it traveled in that time of flight, which is determined by its initial momentum when it was released from the trap. Therefore, this technique measures the momentum and not the position distribution of the atoms. 

@@

\par We first corrected each image for recoil induced detuning using the simulation described above. We simulated a look-up table of predicted optical depths as a function of atomic column density and probe intensity. From our two absorption images we knew the probe intensity and optical depth for each pixel. We then used the simulated look-up table in an inverted fashion to deduce the atomic column density. An example of the effect of this correction procedure in shown in Figure \ref{fig:SampleCorrection}.  To improve the signal and compensate for our shot to shot number fluctuations, we took 15 equivalent images for each data point and, after correcting, added them together.

@@

\caption{An example of our absorption image after 6.784ms time of flight. The 1-D lattice is along the horizontal direction at the center of our images. The two large clouds on the left and right are the atoms in the $\pm 2k_r$ momentum orders that passed through each other unscattered. The smaller cloud in the center is the atoms that remained in the lowest band of the lattice after pulsing, and thus obtained no momentum. The thin spread of atoms around these clouds is the atoms that underwent scattering.   This image was taken coming from below the Feshbach resonance at 20.07 mT. a. Raw optical density b. Optical density of the same image after correcting for recoil induced detuning}  

@@

\par Once we acquired scattering images at a sufficient number of field values, we wanted to extract the fraction of atoms that underwent a single scattering event as a function of the bias field. We focused our attention on single-scattered atoms because, once an atom has scattered multiple times it is hard to determine precisely how many times it scattered, and therefore hard to deduce the scattering cross section.  An atom that underwent a single scattering event, however, is easily identified, as two atoms that scatter elastically will keep the same amplitude of momentum in an arbitrary direction. Therefore, an atom traveling at $2k_r$ to the right that collides with an atom traveling at $2k_r$ to the left will acquire a momentum of $2k_r$ in some direction, and in a time of flight image such atoms will lie in a spherical fashion around the center of mass -  the scattering halo, pictured in Figure \ref{fig:halo}a. 
	\subfigure[]{\includegraphics[scale=0.25]{Picture12.pdf}}
	\subfigure[]{\includegraphics{figure12b.pdf}}
\caption{a. Our experimental setup. After time of flight, the two clouds traveling in opposite $x$ directions have separated and the atoms that underwent a single scattering event were evenly distributed in a scattering halo around the unscattered clouds. An absorption image captured the shadow the clouds cast onto a 2-d plane. b. Inverse Abel transformed image. The atoms within the Fermi radius $k_F$ of each unscattered cloud center are in the unscattered region and counted towards the total unscattered number. The atoms outside the radius $k_r-k_F$ but inside $k_r+k_F$ but outside the unscattered region are counted towards the number of single scattered atoms.   }  
\label{fig:halo}
\par Absorption images captured the integrated column density along $z$, a projected 2-d atomic distribution. In order to extract the radial dependence of the 3-d distribution from the 2-d image, we performed a standard inverse Abel transform. The inverse Abel transform assumes a cylindrical symmetry, which was present in our case, with the axis of symmetry along $x$, defined by the lattice. We thus obtained the atomic distribution as a function of $r$, the radial distance from the scattering center, and $\theta$, the angle between $r$ and symmetry axis $x$, integrated over $\phi$, the azimuthal angle around the $x$ axis. 
\par We then extracted the number of atoms that went through a single scattering event $N_{scat}$, as a fraction of the total atom number $N_{tot}$, for each bias magnetic field value from the transformed images, as shown in figure \ref{fig:halo}b. We obtained the unscattered atom number by counting the number of atoms in the two unscattered clouds. We obtained the number of atoms that underwent a single scattering event by counting the number of atoms outside the Fermi radius of the unscattered clouds, but inside the arc created by rotating the Fermi momentum $k_F$ radius around the original center of the cloud. The total atom number in the image was the sum of those two.
\par We then used our data to deduce the resonant field value $B_0$ and width of the resonance  $\Delta$, the parameters in eq. [\ref{feshbachEq}]. To do this, we fit our data to a model. Since we were in the low energy regime, the scattering cross-section was given by $\sigma=4\pi a^2$. 

@@

\frac{N_{scat}}{N_{tot}^2}=\tilde{a}_{bg}^2\left(1-\frac{\Delta}{B-B_0}\right)^2 + C.
\label{eq:fit}

@@

\par We found that our imaging noise skewed towards the positive, giving rise to a small background offset. We accounted for this in our fit by including a constant offset parameter $C$. 

@@

Our final data is presented in figure \ref{fig:fittedFractions}. The red line is a best fit of the model given in eq. [\ref{eq:fit}]. The fit parameters we extracted are $\tilde{a}_{bg}=1.48e-3$, $\Delta = 1.0(1)$ mT, $B_0 = 20.247(2)$ mT, and $C=9.00e-5$. The error bars on the fitted data were obtained solely from photon shot noise of both absorption images propagated through our analysis.
The accepted values for the $^{40}K$ s-wave Feshbach resonance for the  $\ket{9/2,-9/2}$ and $\ket{9/2,-7/2}$ states are $B_0=20.210(7)$ mT and $\Delta=0.78(6)$ mT. The difference in our measurement may be a result of scattering with the center cloud that recieved no momentum kick, a process that was not taken into account by our analysis. 

@@

\caption{Normalized scattered population plotted versus bias field B. Green dots represent data taken coming from below the resonance, and blue dots represent the data taken coming from above the resonance. Red is the line of best fit, where data coming from above the resonance was used above the resonance and data coming from below the resonance was used below the resonance to create the fit. The regime where the scattering length is likely large enough for the atoms to behave hydrodynamically is shaded in gray. Data points in that region were not used in the fit, as there the assumption $\sigma\rho\ll1$, where $\rho$ is the atom number per unit area, is no longer valid. Values for the resonant field $B_0$ from literature and as found in this work are indicated.    }  

@@

We studied the effects of recoil-induced detuning effects on absorption images and found an optimal imaging time of ~40\us{} for $^{40}K$ atoms for noise minimization after corrections. We use these results to observe s-wave scattering halos of the Fermi gas around the Feshbach resonance and directly verify the resonance location and width. Our analysis can be used in any absorption imaging applications where signal to noise minimization is critical. We performed a new kind of measurement of the resonant magnetic field and with of a Feshbach resonance. 
